---
title: "03-Get-data"
author: "Jelmer"
date: "24-1-2021"
output:
  pdf_document: default
  html_document: default
params:
  data1: dev_stage
  data2: genotype
  geo_dataset: GSE150646
---

## Get the first data of the experiment

With the functions below the data from the selected experiment will be downloaded. The first steps are to collect the data of the experiment and the supplementary files (supp_files). Inside the supp_files are the raw data of the experiment, a supp_file can be any kind of file. For our project we've focused on CSV files only.

Before we can download the files we need to make a directory, the function will be stored here. After that we need to make sure the program knows what we want to work with.
We can do this with parameters (params), you can find them at YAML of the Rmarkdown documents. The used params are geo_dataset (the number of the experiemnt from NCBI_GEO), params data1 and data2 will be explained later.
When you want to analyse several experiments at the same time R needs to split the experiments, so it won't read the params as 1 experiment. We do this with the first code chunk, in this chunk, folders are also made for the different experiments.

```{r }
# Make a directory
# create_dir <- function(path) {
 # 
#}

# When there are more params the (function) "datasets" will split them
datasets <- strsplit(params$geo_dataset, split = " ") %>% unlist
# With full_paths each parameter will be 
full_paths <- file.path(here::here(), datasets)

# Make maps for the different experiments inside geo_dataset
purrr::walk(
  full_paths,
  dir.create)
```

Now the program knows with what kind of data we want to work with, we can download the first set.
We do this with the function “map2” the .x tells what data to look for and .y where to put it. The function getGEO does all the magic to get the data. You might notice at .y we used full_paths, we made this directory in the last code chunk, it gives the location of the folder for each experiment.

```{r}
# Download the data 
data <- map2(
  .x  = datasets,
  .y = full_paths,
  getGEO,
  GSEMatrix = TRUE,
  filename = NULL
)
```

With the first experimental data downloaded, it’s time to get the supp_files. We use the same map2 function from before, but now with the function “getGEOSuppFiles”. This will get the supp_files from the selected experiment.

```{r}
#Get supp files
data_supp <- map2(
  .x  = datasets,
  .y = full_paths,
  getGEOSuppFiles,
  makeDirectory = FALSE, 
  #baseDir = getwd(),
  fetch_files = TRUE, 
  filter_regex = NULL
)
```

Now the experimental data is downloaded (supp_files), we need to unpack the data. Supp_files are often zipped or tarred, our next job is to unzip and or untar these files. First we’ll explain the untarring of the files and after that the unzipping of the data.

