---
title: "06-preSE"
author: "Jelmer"
date: "25-1-2021"
output: html_document
params:
  data1: dev_stage
  data2: genotype
  geo_dataset: GSE150646
---

## Make the data ready for SummarizedExperiment and DESeq2

### Genarating a summarized experiment for CSV

The first steps in making a summarized experiment are getting the phenodata and metadata.
Before we do that we first check if we still have the correct data selected. 
The glimpse(datasets) should give the same output as your selected experiment (params geo_dataset) in this case that is GSE150646.

After this quick check we want to know how to get to our data and make this easier. We see that that the data is saved in a folder inside a folder. To get to the data we need to do the same function twice, that’s why we called the first one 'pre'. Now we can easily ge our data with gse_csv.


```{r}
# We look into the datasets to check if we have selected the experiments we want to work with. 
glimpse(datasets)
glimpse(data)

# To get to the data we need to do the same function twice, thats why we called the first one 'pre'
pre_gse_csv <- data[[1]]
gse_csv <- pre_gse_csv[[1]]
glimpse(gse_csv)
```


Now we know that we’ve the correct data and can easily access them we’ll make the phenodata and metadata. We only need 2 functions to do this, pData and experimentData.
```{r}
# Obtaining the phenotypic data, this contains all the info regarding each sample. 
phenodata_csv <- pData(gse_csv) %>% as_tibble
class(phenodata_csv)
# Make the metadata
metadata_csv <- experimentData(gse_csv)
#class(metadata_csv)
```

Now we need to “clean up” the phenodata a little bit before we can continue making a summarizedExperiment. When we look at the phenodata (with head) we see some weird names like “characteristics_ch1.1” and “characteristics_ch1.2”, these are the variables of the experiments. When you read “characteristics_ch1.1” you’ve no clue what the experiment is about, so you can change the name (for example) to gender/dev_stage/genotype etc. You can change the names with the params data1 and data2 at the start of this document. The place of “characteristics_ch1.1” can change from experiment to experiment, we haven’t find a way to automate this yet.
```{r}
# To look at what columm (number) the characteristics are, this can change per research, we haven't automated this yet.
names(phenodata_csv)
# Het benoemen van experimentele waarde
colnames(phenodata_csv)[11] <- params$data1
colnames(phenodata_csv)[12] <- params$data2


names(phenodata_csv)
```

Further change of the phnodata, needed to make Summarized Experiment.
```{r}
#
phenodata_csv <- phenodata_csv[2:41]
samplenames_csv<- phenodata_csv[1]
```

