---
title: "07-last steps SE"
author: "Jelmer"
date: "25-1-2021"
output: html_document
params:
  data1: dev_stage
  data2: genotype
  geo_dataset: GSE150646
---

### Last steps before the SummarizedExperiment

Now we can load the CSV_file, you’ll see the first few counts inside the file. Check if the data file is a “data.frame".

```{r}
# Load the raw data from the csv data
load_csv_data <- list.files(path = here::here(params$geo_dataset), full.names = TRUE, pattern = ".csv")
csv_data <- read.csv(load_csv_data)

# The first few counts of the raw data
head(csv_data)

# Check if the csv_data is a data.frame
class(csv_data)
```

In raw_counts_csv_data we load in the raw counts of the csv file, minus the first column because these are names of the samples. Also note that it can be more or less columns in different documents. In the next step we change the name of the columns to make them more understandable. The last step is to change the data.frame into a data.matrix for SummarizedExperiment.
```{r}
# In the next part we want to arrange the names. 

# Isolate the raw counts. 
raw_counts_csv_data <- csv_data[2:21]

#head(raw_counts_csv_data)

# Make the rowdata_csv
rowdata_csv_data <- raw_counts_csv_data$`Gene symbol`

# The raw counts are used to create the colnames. 
colnames(raw_counts_csv_data) <-samplenames_csv[[1]]
head(raw_counts_csv_data)

# A matrix is created. 
raw_counts_csv_data_matrix <- data.matrix(raw_counts_csv_data)
rownames(raw_counts_csv_data_matrix) <- rowdata_csv_data
#head(raw_counts_csv_data_matrix)
```

