---
title: "10-visualization"
author: "Jelmer"
date: "25-1-2021"
output: html_document
params:
  data1: dev_stage
  data2: genotype
  geo_dataset: GSE150646
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### The visualization of the dataset 

The code below will plot a histogram, this histogram will show the distribution of the p-values calculated by the DESeq2 analyze. 
```{r}
ggplot(as(res_csv_data, "data.frame"), aes(x = pvalue)) +
  geom_histogram()
```

This code will create a volcano plot, a type of scatterplot, that will help by quickly identifying changes in the dataset. 
```{r}
DESeq2::plotMA(resLFC_csv_data, ylim=c(-2,2))
```

Here a Principal Component Analysis (PCA) is created, a PCA is used to observe outliers, jumps, clusters and trends. 
```{r}
ddsDE_rlog_csv_data <- ddsDE_csv_data %>%
  rlogTransformation() 
ddsDE_rlog_csv_data %>%
  DESeq2::plotPCA(intgroup = c(params$data1 , params$data2)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "red")
```

The code below creates a heatmap, this shows the extent of a phenomenon of the top 20 of up and down regulated genes as colour in two dimensions. 
```{r}
matrix_results_csv <- assay(ddsDE_rlog_csv_data)
ind <- matrix_results_csv %>%
  rowMeans() %>%
  order(decreasing = TRUE) 
top20 <- matrix_results_csv[ind[1:20],]
annot_df <- as.data.frame(colData(ddsDE_csv_data)[,c(params$data1 , params$data2)])
heatmap <- pheatmap(top20,
                    scale = "row", 
                    annotation_col = annot_df)
heatmap
```

