---
title: "13_Sversie_supp_data_download"
author: "Jelmer en Sanne"
date: "27-1-2021"
output: html_document
params:
  data1: dev_stage
  data2: genotype
  geo_dataset: GSE150646
---



## Downloading supfiles and untar/unzip functions. 
### The working code behind the application. 

The code for downloading the supfiles is really simmilar to the download function. Only it loads the supdata, instead of the entire dataset. 
```{r eval= FALSE}
new_supp_download_fucntion <- function(gse_id_number) {
  
  datasets <- strsplit(gse_id_number, split = " ") %>% unlist
  
  full_paths <- file.path(here::here("data"), datasets)  
  
  data_supp <- map2(
    .x  = datasets,
    .y = full_paths,
    getGEOSuppFiles,
    makeDirectory = FALSE, 
    #baseDir = getwd(),
    fetch_files = TRUE, 
    filter_regex = NULL
  ) 
  return(data_supp)
  
}
```
   
This is the old function that is used to unzip and untar the GSE dataset. 
```{r eval= FALSE}
untar_unzip_function <- function(gse_id_number){
  
  #list files that are in the supp files directory 
  files <- list.files(here::here("data", gse_id_number), full.names = TRUE)
  
  #look for .tar extension and use this value to subset the list
  tar_files <- str_subset(files, ".tar")
  
  #loop through the list and untar the files
  for (x in tar_files){
    untar(x, exdir = file.path(here::here("data", gse_id_number)))
  }
  
  #remove tar files because we are done with them
  file.remove(tar_files)
  
  #look again at the files that are now in the directory and list them
  files <- list.files(here::here("data", gse_id_number), full.names = TRUE)
  
  #determine what files are zipped 
  gz_files <- str_subset(files, ".gz")
  
  #loop through the list and unzip the files
  for (x in gz_files){
    gunzip(x)
  }
  
  filenames <- list.files(here::here("data", gse_id_number))
return(filenames)
}
```

The new unzip function is created. We used this function to unzip the CSV data files.    
```{r eval= FALSE}
new_unzip_function <- function(gse_id_number) {
  
  datasets <- strsplit(gse_id_number, split = " ") %>% unlist
  
  full_paths <- file.path(here::here("data"), datasets) 
  
  files_zip <- list.files(path = here::here("data", datasets), full.names = TRUE, pattern = ".gz")
  unzipp <- map(
    .x = files_zip,
    .f = GEOquery::gunzip
  )
  
  return(files_zip)
}   
   
```   
   
## Supplementary files download and make assaydata tab 
   
### Shiny   
#### Server.R  

This is the code used in the server to load the supp files and untar and unzip the data. This code isn't working proprely, one line off code is turned off, because we couldn't create a code that can load the supp files of both text files and CSV files. 
```{r eval= FALSE}   
  #Choose what data will be downloaded
    supp_files_csv <- eventReactive(input$action_download_supp_files_csv, {
    new_supp_download_fucntion(gse_id_number())
  })
  
  #Show a list of supp files found in the folder
  
   output$list_of_downloaded_files <- renderTable(supp_files_csv())
 # output$list_of_downloaded_files <- renderTable(supp_files_txt())
  
  #Untar and unzip the dataset. 
  observeEvent(input$action_untar_unzip_supp_files_csv, {
    new_unzip_function(gse_id_number())
  })
  observeEvent(input$action_untar_unzip_supp_files_txt, {
    untar_unzip_function(gse_id_number())
  })
```

#### UI.R
```{r eval= FALSE}
 tabItem(tabName = "supp_files_assay_data",
            fluidRow(
              
              #select what columns to view
              box(
                title = "download the supplementary files",
                width = "10",
                actionButton("action_download_supp_files_csv", label = "download supp files (CSV)", icon("download", lib = "font-awesome")),
                actionButton("action_download_supp_files_tsv", label = "download supp files (TSV)", icon("download", lib = "font-awesome")),
                actionButton("action_download_supp_files_txt", label = "download supp files (txt)", icon("download", lib = "font-awesome")),
                actionButton("action_download_supp_files_etc", label = "download supp files (etc)", icon("download", lib = "font-awesome"))
                
                
                
              ),
              
              #show a list of supp files found in the folder
              box(
                title = "Show a list of supplementary files that have been downloaded.",
                width = "10",
                tableOutput("list_of_downloaded_files")
              ),
            
              #untar and unzip 
              box(
                title = "untar and unzip files",
                width = "10",
                actionButton("action_untar_unzip_supp_files_csv", label = "untar and unzip supp files csv", icon("download", lib = "font-awesome")),
                actionButton("action_untar_unzip_supp_files_txt", label = "untar and unzip supp files txt", icon("download", lib = "font-awesome"))
                
                ),
              
              #show files that have been untarred and unzipped
              box(
                title = "Show a list of supplementary files that have been unzipped.",
                width = "10",
                verbatimTextOutput("list_of_untarred_files")
              ),
              
              
              #Sample selection box, this uses names of files that are in the data --> "GSE###### folder". The boxes that are selected will be saved to input$sample_choices. 
              box(
                title = "Select samples",
                width = "3",
                uiOutput("sample_choices")
              ),
              
              fluidRow(
                box(
                  width = 10,
                  "Is the sample selection box empty? This is likely due to the folder being empty at the time the GSE number was entered. To fix this simply go to the 'download dataset' tab and click another time on the 'download this GSE dataset' button. 
                  The sample selection boxes should now appear. To fix this problem some sort of 'refresh' button needs to be coded but this is not very straight-forward when submitButtons can not be used."
                ))
              
              
              
              ))
```          


